---
title: "BOP surveys"
author: "Ellie Knott"
date: "2026-09-18"
output: html_document
---

```{r}
require(tidyverse)
require(haven)
require(dplyr)
require(lubridate)
require(forcats)
require(modelsummary)
```

```{r}
# loading data
bop_sep_2025 <- read_sav("Project Data/BOP/DB_BOP_IPP_09_2025_1.SAV")
bop_oct_2024 <- read_sav("Project Data/BOP/BOP_2024.10.SAV")
bop_aug_2023 <- read_sav("Project Data/BOP/MD_BOP_August_2023_cleaned.sav")
bop_nov_2022 <- read_sav("Project Data/BOP/MD_BOP_IPP_Noiembrie_2022_cleaned.sav") 
bop_feb_2021 <- read_sav("Project Data/BOP/2021.02.sav")
bop_oct_2020 <- read_sav("Project Data/BOP/2020.10.sav")
bop_jun_2020 <- read_sav("Project Data/BOP/2020.06.sav")
bop_dec_2019 <- read_sav("Project Data/BOP/2019.12.sav")
bop_nov_2018 <- read_sav("Project Data/BOP/2018.11.sav")
bop_may_2018 <- read_sav("Project Data/BOP/2018.05.sav")
bop_nov_2017 <- read_sav("Project Data/BOP/2017.11.sav")
```

```{r}
waves <- c(bop_nov_2017, bop_may_2018, bop_nov_2018, bop_dec_2019, bop_jun_2020, bop_oct_2020, bop_feb_2021,
              bop_nov_2022, bop_aug_2023, bop_oct_2024, bop_sep_2025)
```


# Aligning, cleaning, merging
```{r}
fix_cyrillic_names <- function(x) {
  cyr <- c("А","В","Е","К","М","Н","О","Р","С","Т","Х","а","е","о","р","с","х","у")
  lat <- c("A","B","E","K","M","H","O","P","C","T","X","a","e","o","p","c","x","y")
  for (i in seq_along(cyr)) {
    x <- gsub(cyr[i], lat[i], x, fixed = TRUE)
  }
  x
}

names(bop_panel) <- fix_cyrillic_names(names(bop_aug_2023))
```


```{r}
attr(bop_sep_2025$D6_3, "labels")
attr(bop_oct_2024$E_3, "labels")
attr(bop_aug_2023$D17_2, "labels")
attr(bop_nov_2022$D17_1, "labels")
attr(bop_feb_2021$varimax, "labels")
attr(bop_oct_2020$B6_1, "labels")
attr(bop_jun_2020$socioecon, "labels")
attr(bop_dec_2019$B6_1, "labels")
attr(bop_nov_2018$C5_9, "labels")
attr(bop_may_2018$B6_01, "labels")
attr(bop_nov_2017$B6_1, "labels")
attr(bop_nov_2017$D13, "label")
```

```{r}
# adding date
bop_sep_2025$date <- dmy("01-09-2025")
bop_oct_2024$date <- dmy("01-10-2024")
bop_aug_2023$date <- dmy("01-08-2023")
bop_nov_2022$date <- dmy("01-11-2022")
bop_feb_2021$date <- dmy("01-02-2021") 
bop_oct_2020$date <- dmy("01-10-2020")
bop_jun_2020$date <- dmy("01-06-2020")
bop_dec_2019$date <- dmy("01-12-2019")
bop_nov_2018$date <- dmy("01-11-2018")
bop_may_2018$date <- dmy("01-05-2018")
bop_nov_2017$date <- dmy("01-11-2017")
```

```{r}
bop_sep_2025$Weights <- bop_sep_2025$W
bop_oct_2024$Weights <- bop_oct_2024$W
bop_aug_2023$Weights <- bop_aug_2023$W1
bop_nov_2022$Weights <- bop_nov_2022$weight_agesex
bop_feb_2021$Weights <- bop_feb_2021$Weights
bop_oct_2020$Weights <- bop_oct_2020$WEGHT
bop_jun_2020$Weights <- bop_jun_2020$weight_age_sex2
bop_dec_2019$Weights <- bop_dec_2019$w
bop_nov_2018$Weights <- bop_nov_2018$weight_age_sex2
bop_may_2018$Weights <- bop_may_2018$weight
bop_nov_2017$Weights <- bop_nov_2017$weight_age_sex2
```

## Demographics

```{r}
# Gender
gender_labels <- c("Man", "Woman")

bop_sep_2025$Gender <- factor(bop_sep_2025$Genul, labels = gender_labels)
bop_oct_2024$Gender <- factor(bop_oct_2024$S_8, labels = gender_labels)
bop_aug_2023$Gender <- factor(bop_aug_2023$Genul, labels = gender_labels)
bop_nov_2022$Gender <- factor(bop_nov_2022$Genul, labels = gender_labels)
bop_feb_2021$Gender <- factor(bop_feb_2021$Genul, labels = gender_labels)
bop_oct_2020$Gender <- factor(bop_oct_2020$Genul, labels = gender_labels)
bop_jun_2020$Gender <- factor(bop_jun_2020$genukl, labels = gender_labels)
bop_dec_2019$Gender <- factor(bop_dec_2019$S1, labels = gender_labels)
bop_nov_2018$Gender <- factor(bop_nov_2018$Genul, labels = gender_labels)
bop_may_2018$Gender <- factor(bop_may_2018$S1, labels = gender_labels)
bop_nov_2017$Gender <- factor(bop_nov_2017$S1, labels = gender_labels)
```

```{r}
# Age

bop_sep_2025$Age <- bop_sep_2025$G2
bop_oct_2024$Age <- bop_oct_2024$S_2
bop_aug_2023$Age <- bop_aug_2023$S2
bop_nov_2022$Age <- bop_nov_2022$S2
bop_feb_2021$Age <- bop_feb_2021$S2
bop_oct_2020$Age <- bop_oct_2020$S2
bop_jun_2020$Age <- bop_jun_2020$S2
bop_dec_2019$Age <- bop_dec_2019$S2
bop_nov_2018$Age <- bop_nov_2018$S2
bop_may_2018$Age <- bop_may_2018$S2
bop_nov_2017$Age <- bop_nov_2017$S2

# As factor
age_levels <- c("18-29", "30-44", "45-59", "60-74", "75+")

bop_sep_2025$Age.f <- cut(bop_sep_2025$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_oct_2024$Age.f <- cut(bop_oct_2024$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_aug_2023$Age.f <- cut(bop_aug_2023$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_nov_2022$Age.f <- cut(bop_nov_2022$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_feb_2021$Age.f <- cut(bop_feb_2021$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_oct_2020$Age.f <- cut(bop_oct_2020$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_jun_2020$Age.f <- cut(bop_jun_2020$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_dec_2019$Age.f <- cut(bop_dec_2019$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_nov_2018$Age.f <- cut(bop_nov_2018$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_may_2018$Age.f <- cut(bop_may_2018$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
bop_nov_2017$Age.f <- cut(bop_nov_2017$Age, breaks = c(17, 29, 44, 59, 74, Inf), labels = age_levels)
```

```{r}
# Ethnicity
bop_sep_2025$Ethnicity <- factor(bop_sep_2025$S5, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Other", "Bulgarian", "Other", "Other", "NR"))
bop_oct_2024$Ethnicity <- factor(bop_oct_2024$S_3, levels = c(1:11, 99), labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "Other", "Other", "Other", "Other", "NR"))
bop_aug_2023$Ethnicity <- factor(bop_aug_2023$S5, levels = 1:8, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_nov_2022$Ethnicity <- factor(bop_nov_2022$S7, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_feb_2021$Ethnicity <- factor(bop_feb_2021$S6, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_oct_2020$Ethnicity <- factor(bop_oct_2020$S3, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_jun_2020$Ethnicity <- factor(bop_jun_2020$S3, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_dec_2019$Ethnicity <- factor(bop_dec_2019$S3, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_nov_2018$Ethnicity <- factor(bop_nov_2018$S3, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_may_2018$Ethnicity <- factor(bop_may_2018$S3, levels = 1:8, labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))
bop_nov_2017$Ethnicity <- factor(bop_nov_2017$S3, levels = c(1:7,99), labels = c("Moldovan", "Romanian", "Russian", "Ukrainian", "Gagauz", "Bulgarian", "Other", "NR"))

## unfortunately, inconsistent coding means we have to get rid of those identifying as Roma
```

```{r}
# Language of questionnaire
questionnaire_labels <- c("Romanian", "Russian")

bop_sep_2025$Language_Questionnaire <- factor(bop_sep_2025$Lang, labels = questionnaire_labels)
bop_oct_2024$Language_Questionnaire <- factor(bop_oct_2024$S_12, labels = questionnaire_labels)
bop_aug_2023$Language_Questionnaire <- factor(bop_aug_2023$Lang, labels = questionnaire_labels)
bop_nov_2022$Language_Questionnaire <- factor(bop_nov_2022$Lang, labels = questionnaire_labels)
bop_feb_2021$Language_Questionnaire <- factor(bop_feb_2021$Lang, labels = questionnaire_labels)
bop_oct_2020$Language_Questionnaire <- factor(bop_oct_2020$S12, labels = questionnaire_labels)
bop_jun_2020$Language_Questionnaire <- factor(bop_jun_2020$S11, labels = questionnaire_labels)
bop_dec_2019$Language_Questionnaire <- factor(bop_dec_2019$S18, labels = questionnaire_labels)
bop_nov_2018$Language_Questionnaire <- factor(bop_nov_2018$S18, labels = questionnaire_labels)
bop_may_2018$Language_Questionnaire <- factor(bop_may_2018$Limba, labels = questionnaire_labels)
bop_nov_2017$Language_Questionnaire <- factor(bop_nov_2017$S18, labels = questionnaire_labels)
```

```{r}
# rural / urban ecology
ecology_labels <- c("urban", "rural")

bop_sep_2025$Ecology <- factor(bop_sep_2025$mediu, labels = ecology_labels)
bop_oct_2024$Ecology <- NA
bop_aug_2023$Ecology <- factor(bop_aug_2023$mediu, labels = ecology_labels)
bop_nov_2022$Ecology <- factor(bop_nov_2022$mediu, labels = ecology_labels)
bop_feb_2021$Ecology <- factor(bop_feb_2021$Mediu, labels = ecology_labels)
bop_oct_2020$Ecology <- factor(bop_oct_2020$mediu, labels = ecology_labels)
bop_jun_2020$Ecology <- factor(bop_jun_2020$mediu, labels = ecology_labels)
bop_dec_2019$Ecology <- factor(bop_dec_2019$mediu, labels = ecology_labels)
bop_nov_2018$Ecology <- factor(bop_nov_2018$mediu, labels = ecology_labels)
bop_may_2018$Ecology <- factor(bop_may_2018$S18, labels = ecology_labels)
bop_nov_2017$Ecology <- factor(bop_nov_2017$mediul, labels = ecology_labels)
```

```{r}
# Education
education_labels <- c("low", "medium", "high")

bop_sep_2025$Education <- factor(bop_sep_2025$studii, labels = education_labels)
bop_oct_2024$Education <- factor(bop_oct_2024$S_4, labels = c("low", "low", "medium", "medium", "medium", "high", "high", "high", "high", "high"))
bop_aug_2023$Education <- factor(bop_aug_2023$studii, labels = education_labels)
bop_nov_2022$Education <- factor(bop_nov_2022$studii, labels = education_labels)
bop_feb_2021$Education <- factor(bop_feb_2021$studii, labels = education_labels)
bop_oct_2020$Education <- factor(bop_oct_2020$educatie, labels = education_labels)
bop_jun_2020$Education <- factor(bop_jun_2020$studii, labels = education_labels)
bop_dec_2019$Education <- factor(bop_dec_2019$S4, labels = c("low", "low", "medium", "medium", "medium", "high", "high", "high", "high", "DK", "NR"))
bop_nov_2018$Education <- factor(bop_nov_2018$studii, labels = education_labels)
bop_may_2018$Education <- factor(bop_may_2018$EDUCATIE, labels = education_labels)
bop_nov_2017$Education <- factor(bop_nov_2017$studiile, labels = education_labels)
```

```{r}
# Region
regions1 <- c("Anenii Noi", "Balti", "Basarabeasca", "Briceni", "Cahul", "Calarasi", "Cantemir", "Causeni", "Chisinau", "Cimislia", "Criuleni", "Donduseni", "Drochia", "Dubasari", "Edinet", "Falesti", "Floresti", "Glodeni", "Hincesti", "Ialoveni", "Leova", "Nisporeni", "Ocnita", "Orhei", "Rezina", "Riscani", "Singerei", "Soldanesti", "Soroca", "Stefan Voda", "Straseni", "Taraclia", "Telenesti", "Ungheni", "Gagauzia")
regions2 <- c("Chisinau", "Anenii Noi", "Balti", "Basarabeasca", "Briceni", "Cahul", "Calarasi", "Cantemir", "Causeni", "Cimislia", "Criuleni", "Donduseni", "Drochia", "Dubasari", "Edinet", "Falesti", "Floresti", "Glodeni", "Hincesti", "Ialoveni", "Leova", "Nisporeni", "Ocnita", "Orhei", "Rezina", "Riscani", "Singerei", "Soldanesti", "Soroca", "Stefan Voda", "Straseni", "Taraclia", "Telenesti", "Ungheni", "Gagauzia")

bop_sep_2025$Region <- factor(bop_sep_2025$mun, labels = regions2)
bop_oct_2024$Region <- NA
bop_aug_2023$Region <- factor(bop_aug_2023$mun, labels = regions2)
bop_nov_2022$Region <- factor(bop_nov_2022$mun, labels = regions2)
bop_feb_2021$Region <- factor(bop_feb_2021$mun, labels = regions2)
bop_oct_2020$Region <- factor(bop_oct_2020$S15, labels = regions2)
bop_jun_2020$Region <- factor(bop_jun_2020$S10, labels = regions2)
bop_dec_2019$Region <- NA
bop_nov_2018$Region <- factor(bop_nov_2018$S21, labels = regions2)
bop_may_2018$Region <- factor(bop_may_2018$S20, labels = regions1)
bop_nov_2017$Region <- factor(bop_nov_2017$S21, labels = regions2)
```

```{r}
# macro regions
macro_region <- c("North", "Centre", "Chisinau", "South", "Gagauzia")
bop_oct_2024$Macro_Region <- NA
bop_dec_2019$Macro_Region <- factor(bop_dec_2019$reg, labels = macro_region)
bop_sep_2025$Macro_Region <- fct_collapse(bop_sep_2025$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_aug_2023$Macro_Region <- fct_collapse(bop_aug_2023$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_nov_2022$Macro_Region <- fct_collapse(bop_nov_2022$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_feb_2021$Macro_Region <- fct_collapse(bop_feb_2021$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_oct_2020$Macro_Region <- fct_collapse(bop_oct_2020$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_jun_2020$Macro_Region <- fct_collapse(bop_jun_2020$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_nov_2018$Macro_Region <- fct_collapse(bop_nov_2018$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_may_2018$Macro_Region <- fct_collapse(bop_may_2018$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")

bop_nov_2017$Macro_Region <- fct_collapse(bop_nov_2017$Region,
                             "North"    = c("Balti", "Briceni", "Donduseni", "Drochia", "Edinet",
                                            "Falesti", "Floresti", "Glodeni", "Ocnita", "Riscani",
                                            "Singerei", "Soldanesti", "Soroca"),
                             "Centre"   = c("Anenii Noi", "Calarasi", "Criuleni", "Dubasari",
                                            "Hincesti", "Ialoveni", "Nisporeni", "Orhei", "Rezina",
                                            "Straseni", "Telenesti", "Ungheni"),
                             "South"    = c("Basarabeasca", "Cahul", "Cantemir", "Causeni",
                                            "Cimislia", "Leova", "Stefan Voda", "Taraclia"),
                             "Chisinau" = "Chisinau",
                             "Gagauzia" = "Gagauzia")
```

```{r}
#socioeconomic
socioeconomic_labels <- c("low", "medium", "high")

bop_sep_2025$Socioeconomic <- factor(bop_sep_2025$Varimax, labels = socioeconomic_labels)
bop_oct_2024$Socioeconomic <- NA
bop_aug_2023$Socioeconomic <- factor(bop_aug_2023$Niv.socio, labels = socioeconomic_labels)
bop_nov_2022$Socioeconomic <- factor(bop_nov_2022$varimax, labels = socioeconomic_labels)
bop_feb_2021$Socioeconomic <- factor(bop_feb_2021$varimax, labels = socioeconomic_labels)
bop_oct_2020$Socioeconomic <- factor(bop_oct_2020$varimax, labels = socioeconomic_labels)
bop_jun_2020$Socioeconomic <- factor(bop_jun_2020$socioecon, labels = socioeconomic_labels)
bop_dec_2019$Socioeconomic <- factor(bop_dec_2019$st, labels = socioeconomic_labels)
bop_nov_2018$Socioeconomic <- factor(bop_nov_2018$Varimax, labels = socioeconomic_labels)
bop_may_2018$Socioeconomic <- factor(bop_may_2018$N_S_E, labels = socioeconomic_labels)
bop_nov_2017$Socioeconomic <- factor(bop_nov_2017$varimax, labels = socioeconomic_labels)
```

## Moldova's Direction
```{r}
direction_labels <- c("Good", "Bad", "DK", "NR", NA)

bop_nov_2017$direction <- factor(bop_nov_2017$A1, levels = c(1, 2, 98, 99, 5), labels = direction_labels)
bop_may_2018$direction <- factor(bop_may_2018$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_nov_2018$direction <- factor(bop_nov_2018$A1, levels = c(1, 2, 98, 99, 5), labels = direction_labels)
bop_dec_2019$direction <- factor(bop_dec_2019$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_jun_2020$direction <- factor(bop_jun_2020$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_oct_2020$direction <- factor(bop_oct_2020$A1, levels = c(1, 2, 98, 99, 5), labels = direction_labels)
bop_feb_2021$direction <- factor(bop_feb_2021$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_nov_2022$direction <- factor(bop_nov_2022$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_aug_2023$direction <- factor(bop_aug_2023$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_oct_2024$direction <- factor(bop_oct_2024$A_1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
bop_sep_2025$direction <- factor(bop_sep_2025$A1, levels = c(1, 2, 3, 4, 5), labels = direction_labels)
```

## Problems
```{r}
# First 
bop_sep_2025$First_Problem <- factor(bop_sep_2025$D6_1, levels = 1:16, labels =  c("Internal order", "Peace and understanding among people", "Infrastructure", "Economy", "Standard of living", "Education, science, culture", "Social services", "Social cohesion", "Spiritual development", "Territorial integrity (Transnistria)", "Cooperation with IOs", "Corruption", "Other", "DK", "NR", "NR"))
bop_oct_2024$First_Problem <- factor(bop_oct_2024$A_7_1, levels = c(1:11, 98, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_aug_2023$First_Problem <- factor(bop_aug_2023$D17_1, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_nov_2022$First_Problem <- factor(bop_nov_2022$D17_1, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_oct_2020$First_Problem <- factor(bop_oct_2020$B6_1, levels = c(1:11, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_dec_2019$First_Problem <- factor(bop_dec_2019$B6_1, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_may_2018$First_Problem <- factor(bop_may_2018$B6_01, levels = c(1:12), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_nov_2017$First_Problem <- factor(bop_nov_2017$B6_1, levels = c(1:11, 100), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
```

```{r}
# Second
bop_sep_2025$Second_Problem <- factor(bop_sep_2025$D6_2, levels = 1:16, labels =  c("Internal order", "Peace and understanding among people", "Infrastructure", "Economy", "Standard of living", "Education, science, culture", "Social services", "Social cohesion", "Spiritual development", "Territorial integrity (Transnistria)", "Cooperation with IOs", "Corruption", "Other", "DK", "NR", "NR"))
bop_oct_2024$Second_Problem <- factor(bop_oct_2024$A_7_2, levels = c(1:11, 98, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_aug_2023$Second_Problem <- factor(bop_aug_2023$D17_2, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_nov_2022$Second_Problem <- factor(bop_nov_2022$D17_2, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_oct_2020$Second_Problem <- factor(bop_oct_2020$B6_2, levels = c(1:11, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_dec_2019$Second_Problem <- factor(bop_dec_2019$B6_2, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_may_2018$Second_Problem <- factor(bop_may_2018$B6_02, levels = c(1:12), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_nov_2017$Second_Problem <- factor(bop_nov_2017$B6_2, levels = c(1:11, 100), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
```

```{r}
# Third
bop_sep_2025$Third_Problem <- factor(bop_sep_2025$D6_3, levels = 1:16, labels =  c("Internal order", "Peace and understanding among people", "Infrastructure", "Economy", "Standard of living", "Education, science, culture", "Social services", "Social cohesion", "Spiritual development", "Territorial integrity (Transnistria)", "Cooperation with IOs", "Corruption", "Other", "DK", "NR", "NR"))
bop_oct_2024$Third_Problem <- factor(bop_oct_2024$A_7_3, levels = c(1:11, 98, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_aug_2023$Third_Problem <- factor(bop_aug_2023$D17_3, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_nov_2022$Third_Problem <- factor(bop_nov_2022$D17_3, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_oct_2020$Third_Problem <- factor(bop_oct_2020$B6_3, levels = c(1:11, 99), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_dec_2019$Third_Problem <- factor(bop_dec_2019$B6_3, levels = c(1:13), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK", "NR"))
bop_may_2018$Third_Problem <- factor(bop_may_2018$B6_03, levels = c(1:12), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
bop_nov_2017$Third_Problem <- factor(bop_nov_2017$B6_3, levels = c(1:11, 100), labels =  c("Internal order", "Peace and understanding among people", "Economy", "Standard of living", "Education, science, culture", "Spiritual development", "Territorial integrity (Transnistria)", "Independent Judiciary", "Cooperation with IOs", "Corruption", "Other", "DK"))
```

## Media
```{r}
# first important media
bop_sep_2025$Important_Media <- factor(bop_sep_2025$B10_1, labels =  c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "Social media", "DK", "DK", "NR"))
bop_oct_2024$Important_Media <- factor(bop_oct_2024$C_13_O1, levels = c(1:10, 96:99), labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Internet", "Social media", "Social media", "Other", "Other", "None", "DK", "DK", "NR"))
bop_aug_2023$Important_Media <- factor(bop_aug_2023$С3_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_nov_2022$Important_Media <- factor(bop_nov_2022$С3_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_feb_2021$Important_Media <- factor(bop_feb_2021$B10_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_oct_2020$Important_Media <- factor(bop_oct_2020$B11_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "NR"))
bop_jun_2020$Important_Media <- factor(bop_jun_2020$B4_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Internet", "DK", "DK", "NR"))
bop_dec_2019$Important_Media <- factor(bop_dec_2019$B10, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Internet", "None", "DK", "DK", "NR"))
bop_nov_2018$Important_Media <- factor(bop_nov_2018$B10_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Internet", "DK", "DK", "NR"))
bop_may_2018$Important_Media <- factor(bop_may_2018$B10_01, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_nov_2017$Important_Media <- factor(bop_nov_2017$B10_1, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Internet", "DK", "DK", "NR"))

```


```{r}
# second important media
bop_sep_2025$Second_Media <- factor(bop_sep_2025$B10_2, labels =  c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "Social media"))
bop_oct_2024$Second_Media <- factor(bop_oct_2024$C_13_O2, levels = c(1:10, 96:99), labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Internet", "Social media", "Social media", "Other", "Other", "None", "DK", "DK", "NR"))
bop_aug_2023$Second_Media <- factor(bop_aug_2023$C3_2, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_nov_2022$Second_Media <- factor(bop_nov_2022$С3_2, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet"))
bop_feb_2021$Second_Media <- factor(bop_feb_2021$B10_2, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_oct_2020$Second_Media <- factor(bop_oct_2020$B11_2, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_jun_2020$Second_Media <- factor(bop_jun_2020$B4_2, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_dec_2019$Second_Media <- factor(bop_dec_2019$B10a, levels = 1:14, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "DK", "NR"))
bop_nov_2018$Second_Media <- factor(bop_nov_2018$B10_2, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet"))
bop_may_2018$Second_Media <- factor(bop_may_2018$B10_02, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "DK", "NR"))
bop_nov_2017$Second_Media <- factor(bop_nov_2017$B10_2, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet"))
```


```{r}
# most trusted media 
bop_sep_2025$Trust_Media <- factor(bop_sep_2025$B11, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "Social media", "DK", "DK", "NR", "NA"))
bop_oct_2024$Trust_Media <- factor(bop_oct_2024$C_14, levels = c(1:10, 96:99), labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Internet", "Social media", "Social media", "Other", "Other", "None", "DK", "DK", "NR"))
bop_aug_2023$Trust_Media <- factor(bop_aug_2023$С4, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "NR"))
bop_nov_2022$Trust_Media <- factor(bop_nov_2022$С4, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "NR"))
bop_feb_2021$Trust_Media <- factor(bop_feb_2021$B11, levels = 1:13, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "NR"))
bop_oct_2020$Trust_Media <- factor(bop_oct_2020$B12, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "DK", "NR"))
bop_jun_2020$Trust_Media <- factor(bop_jun_2020$B5, levels = 1:14, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "DK", "NR"))
bop_dec_2019$Trust_Media <- factor(bop_dec_2019$B11, levels = 1:14, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "DK", "NR"))
bop_nov_2018$Trust_Media <- factor(bop_nov_2018$B11, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Internet", "None", "DK", "DK", "NR"))
bop_may_2018$Trust_Media <- factor(bop_may_2018$B11, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "DK", "NR"))
bop_nov_2017$Trust_Media <- factor(bop_nov_2017$B11, labels = c("TV", "Newspapers", "Radio", "Family", "Friends and neighbours", "Colleagues", "Hearsay", "Opinion polls", "Local authorities", "Internet", "None", "DK", "DK", "NR"))
```

## Domestic Politics

```{r}
# Sandu
bop_sep_2025$Sandu <- factor(bop_sep_2025$D9_3, levels = 1:4)
bop_oct_2024$Sandu <- factor(bop_oct_2024$D_16_1, levels = 1:4)
bop_aug_2023$Sandu <- factor(bop_aug_2023$D10_2, levels = 1:4)
bop_nov_2022$Sandu <- factor(bop_nov_2022$D11_3, levels = 1:4)
bop_feb_2021$Sandu <- factor(bop_feb_2021$D5_2, levels = 1:4)
bop_oct_2020$Sandu <- factor(bop_oct_2020$C5_6, levels = 1:4)
bop_jun_2020$Sandu <- factor(bop_jun_2020$C4_1, levels = 1:4)
bop_dec_2019$Sandu <- factor(bop_dec_2019$C11_3, levels = 1:4)
bop_nov_2018$Sandu <- factor(bop_nov_2018$C5_9, levels = 1:4)
bop_may_2018$Sandu <- factor(bop_may_2018$C5_07, levels = 1:4)
bop_nov_2017$Sandu <- factor(bop_nov_2017$C5_7, levels = 1:4)
```

```{r}
# Dodon
bop_sep_2025$Dodon <- factor(bop_sep_2025$D9_2, levels = 1:4)
bop_oct_2024$Dodon <- factor(bop_oct_2024$D_16_4, levels = 1:4)
bop_aug_2023$Dodon <- factor(bop_aug_2023$D10_3, levels = 1:4)
bop_nov_2022$Dodon <- factor(bop_nov_2022$D11_4, levels = 1:4)
bop_feb_2021$Dodon <- factor(bop_feb_2021$D5_1, levels = 1:4)
bop_oct_2020$Dodon <- factor(bop_oct_2020$C5_4, levels = 1:4)
bop_jun_2020$Dodon <- factor(bop_jun_2020$C4_2, levels = 1:4)
bop_dec_2019$Dodon <- factor(bop_dec_2019$C11_10, levels = 1:4)
bop_nov_2018$Dodon <- factor(bop_nov_2018$C5_6, levels = 1:4)
bop_may_2018$Dodon <- factor(bop_may_2018$C5_05, levels = 1:4)
bop_nov_2017$Dodon <- factor(bop_nov_2017$C5_5, levels = 1:4)
```

```{r}
# Voronin
bop_sep_2025$Voronin <- factor(bop_sep_2025$D9_10, levels = 1:4)
bop_oct_2024$Voronin <- NA
bop_aug_2023$Voronin <- factor(bop_aug_2023$D10_1, levels = 1:4)
bop_nov_2022$Voronin <- factor(bop_nov_2022$D11_1, levels = 1:4)
bop_feb_2021$Voronin <- factor(bop_feb_2021$D5_6, levels = 1:4)
bop_oct_2020$Voronin <- factor(bop_oct_2020$C5_11, levels = 1:4)
bop_jun_2020$Voronin <- factor(bop_jun_2020$C4_10, levels = 1:4)
bop_dec_2019$Voronin <- factor(bop_dec_2019$C11_8, levels = 1:4)
bop_nov_2018$Voronin <- factor(bop_nov_2018$C5_15, levels = 1:4)
bop_may_2018$Voronin <- factor(bop_may_2018$C5_13, levels = 1:4)
bop_nov_2017$Voronin <- factor(bop_nov_2017$C5_13, levels = 1:4)
```

```{r}
# Sor
bop_sep_2025$Sor <- factor(bop_sep_2025$D9_4, levels = 1:4)
bop_oct_2024$Sor <- NA
bop_aug_2023$Sor <- factor(bop_aug_2023$D10_5, levels = 1:4)
bop_nov_2022$Sor <- factor(bop_nov_2022$D11_6, levels = 1:4)
bop_feb_2021$Sor <- factor(bop_feb_2021$D5_7, levels = 1:4)
bop_oct_2020$Sor <- factor(bop_oct_2020$C5_5, levels = 1:4)
bop_jun_2020$Sor <- factor(bop_jun_2020$C4_9, levels = 1:4)
bop_dec_2019$Sor <- factor(bop_dec_2019$C11_5, levels = 1:4)
bop_nov_2018$Sor <- factor(bop_nov_2018$C5_7, levels = 1:4)
bop_may_2018$Sor <- NA
bop_nov_2017$Sor <- NA
```

```{r}
# PAS
bop_sep_2025$PAS <- factor(bop_sep_2025$D10_1, levels = 1:4)
bop_oct_2024$PAS <- factor(bop_oct_2024$D_5_1, levels = 1:4)
bop_aug_2023$PAS <- factor(bop_aug_2023$D5_2, levels = 1:4)
bop_nov_2022$PAS <- factor(bop_nov_2022$D9_2, levels = 1:4)
bop_feb_2021$PAS <- factor(bop_feb_2021$D6_1, levels = 1:4)
bop_oct_2020$PAS <- factor(bop_oct_2020$C3_1, levels = 1:4)
bop_jun_2020$PAS <- factor(bop_jun_2020$C5_1, levels = 1:4)
bop_dec_2019$PAS <- factor(bop_dec_2019$C9_3, levels = 1:4)
bop_nov_2018$PAS <- factor(bop_nov_2018$C3_1, levels = 1:4)
bop_may_2018$PAS <- factor(bop_may_2018$C3_01, levels = 1:4)
bop_nov_2017$PAS <- factor(bop_nov_2017$C3_1, levels = 1:4)
```

```{r}
# PSRM
bop_sep_2025$PSRM <- NA
bop_oct_2024$PSRM <- factor(bop_oct_2024$D_5_4, levels = 1:4)
bop_aug_2023$PSRM <- factor(bop_aug_2023$D5_1, levels = 1:4)
bop_nov_2022$PSRM <- factor(bop_nov_2022$D9_2, levels = 1:4)
bop_feb_2021$PSRM <- factor(bop_feb_2021$D6_5, levels = 1:4)
bop_oct_2020$PSRM <- factor(bop_oct_2020$C3_4, levels = 1:4)
bop_jun_2020$PSRM <- factor(bop_jun_2020$C5_4, levels = 1:4)
bop_dec_2019$PSRM <- factor(bop_dec_2019$C9_1, levels = 1:4)
bop_nov_2018$PSRM <- factor(bop_nov_2018$C3_10, levels = 1:4)
bop_may_2018$PSRM <- factor(bop_may_2018$C3_10, levels = 1:4)
bop_nov_2017$PSRM <- factor(bop_nov_2017$C3_10, levels = 1:4)
```

## International Politics
```{r}
referendum_labels <- c("For", "Against", "Wouldn't participate", "DK", "NR")

bop_sep_2025$Unification <- factor(bop_sep_2025$H10, labels = referendum_labels)
bop_oct_2024$Unification <- factor(bop_oct_2024$E_6, labels = referendum_labels)
bop_aug_2023$Unification <- factor(bop_aug_2023$E8, labels = referendum_labels)
bop_nov_2022$Unification <- factor(bop_nov_2022$E12, labels = referendum_labels)
bop_feb_2021$Unification <- NA
bop_oct_2020$Unification <- factor(bop_oct_2020$D5, labels = referendum_labels)
bop_jun_2020$Unification <- NA
bop_dec_2019$Unification <- NA
bop_nov_2018$Unification <- NA
bop_may_2018$Unification <- factor(bop_may_2018$D8, labels = referendum_labels)
bop_nov_2017$Unification <- factor(bop_nov_2017$D8, labels = referendum_labels)
```

```{r}
bop_sep_2025$NATO <- factor(bop_sep_2025$H10, labels = referendum_labels)
bop_oct_2024$NATO <- NA
bop_aug_2023$NATO <- factor(bop_aug_2023$E7, labels = referendum_labels)
bop_nov_2022$NATO <- factor(bop_nov_2022$E10, labels = referendum_labels)
bop_feb_2021$NATO <- NA
bop_oct_2020$NATO <- NA
bop_jun_2020$NATO <- NA
bop_dec_2019$NATO <- NA
bop_nov_2018$NATO <- NA
bop_may_2018$NATO <- factor(bop_may_2018$D13, labels = referendum_labels)
bop_nov_2017$NATO <- factor(bop_nov_2017$D13, labels = referendum_labels)
```

```{r}
# EU
bop_sep_2025$EU <- factor(bop_sep_2025$H4, labels = referendum_labels)
bop_oct_2024$EU <- factor(bop_oct_2024$E_3, labels = referendum_labels)
bop_aug_2023$EU <- factor(bop_aug_2023$E2, labels = referendum_labels)
bop_nov_2022$EU <- factor(bop_nov_2022$E5, labels = referendum_labels)
bop_feb_2021$EU <- factor(bop_feb_2021$H7, labels = referendum_labels)
bop_oct_2020$EU <- factor(bop_oct_2020$D2, labels = referendum_labels)
bop_jun_2020$EU <- factor(bop_jun_2020$E2, labels = referendum_labels)
bop_dec_2019$EU <- factor(bop_dec_2019$D2, labels = referendum_labels)
bop_nov_2018$EU <- factor(bop_nov_2018$D2, labels = referendum_labels)
bop_may_2018$EU <- factor(bop_may_2018$D2, labels = referendum_labels)
bop_nov_2017$EU <- factor(bop_nov_2017$D2, labels = referendum_labels)
```

```{r}
# Eurasia

bop_sep_2025$Eurasia <- factor(bop_sep_2025$H5, labels = referendum_labels)
bop_oct_2024$Eurasia <- factor(bop_oct_2024$E_5, labels = referendum_labels)
bop_aug_2023$Eurasia <- factor(bop_aug_2023$E4, labels = referendum_labels)
bop_nov_2022$Eurasia <- factor(bop_nov_2022$E7, labels = referendum_labels)
bop_feb_2021$Eurasia <- factor(bop_feb_2021$H9, labels = referendum_labels)
bop_oct_2020$Eurasia <- factor(bop_oct_2020$D3, labels = referendum_labels)
bop_jun_2020$Eurasia <- factor(bop_jun_2020$E3, labels = referendum_labels)
bop_dec_2019$Eurasia <- NA
bop_nov_2018$Eurasia <- factor(bop_nov_2018$D4, labels = referendum_labels)
bop_may_2018$Eurasia <- factor(bop_may_2018$D4, labels = referendum_labels)
bop_nov_2017$Eurasia <- factor(bop_nov_2017$D4, labels = referendum_labels)
```



```{r}
# EU vs Eurasia
EU_Eurasia_labels <- c("EU", "Eurasia", "Wouldn't participate", "DK", "NR")

bop_sep_2025$EU_vs_Eurasia <- factor(bop_sep_2025$H5, labels = EU_Eurasia_labels)
bop_oct_2024$EU_vs_Eurasia <- NA
bop_aug_2023$EU_vs_Eurasia <- factor(bop_aug_2023$E6, labels = EU_Eurasia_labels)
bop_nov_2022$EU_vs_Eurasia <- factor(bop_nov_2022$E9, labels = EU_Eurasia_labels)
bop_feb_2021$EU_vs_Eurasia <- factor(bop_feb_2021$H11, labels = EU_Eurasia_labels)
bop_oct_2020$EU_vs_Eurasia <- factor(bop_oct_2020$D4, labels = EU_Eurasia_labels)
bop_jun_2020$EU_vs_Eurasia <- factor(bop_jun_2020$E4, labels = EU_Eurasia_labels)
bop_dec_2019$EU_vs_Eurasia <- NA
bop_nov_2018$EU_vs_Eurasia <- factor(bop_nov_2018$D6, labels = EU_Eurasia_labels)
bop_may_2018$EU_vs_Eurasia <- factor(bop_may_2018$D6, labels = EU_Eurasia_labels)
bop_nov_2017$EU_vs_Eurasia <- factor(bop_nov_2017$D6, labels = EU_Eurasia_labels)
```

# Merging
```{r}
# giving each way a name
bop_sep_2025$wave <- "mu"
bop_oct_2024$wave <- "lamba"
bop_aug_2023$wave <- "kappa"
bop_nov_2022$wave <- "iota"
bop_feb_2021$wave <- "theta"
bop_oct_2020$wave <- "eta"
bop_jun_2020$wave <- "zeta"
bop_dec_2019$wave <- "epsilon"
bop_nov_2018$wave <- "delta"
bop_may_2018$wave <- "beta"
bop_nov_2017$wave <- "alpha"
```

```{r}
bop_sep_2025$id <- bop_sep_2025$INTNR
bop_oct_2024$id <- bop_oct_2024$SbjNum
bop_aug_2023$id <- bop_aug_2023$INTNR
bop_nov_2022$id <- bop_nov_2022$INTNR
bop_feb_2021$id <- bop_feb_2021$INTNR
bop_oct_2020$id <- bop_oct_2020$INTNR
bop_jun_2020$id <- bop_jun_2020$INTNR
bop_dec_2019$id <- bop_dec_2019$ID
bop_nov_2018$id <- bop_nov_2018$INTNR
bop_nov_2017$id <- bop_nov_2017$INTNR

bop_may_2018 <- bop_may_2018 %>%
    mutate(id = row_number())
```

```{r}
vars_to_keep <- c("id", "wave", "date", "Age", "Age.f", "Gender", "Ethnicity", "Language_Questionnaire", "Ecology", "Education", "Socioeconomic", "Region", "Macro_Region", "direction", "First_Problem", "Second_Problem", "Third_Problem", "Important_Media", "Second_Media", "Trust_Media", "Sandu", "Dodon", "Voronin", "Sor", "PAS", "PSRM", "Unification", "NATO", "EU", "Eurasia", "EU_vs_Eurasia", "Weights")

bop_panel <- bind_rows(
  bop_sep_2025 %>% select(any_of(vars_to_keep)),
  bop_oct_2024 %>% select(any_of(vars_to_keep)),
  bop_aug_2023 %>% select(any_of(vars_to_keep)),
  bop_nov_2022 %>% select(any_of(vars_to_keep)),
  bop_feb_2021 %>% select(any_of(vars_to_keep)),
  bop_oct_2020 %>% select(any_of(vars_to_keep)),
  bop_jun_2020 %>% select(any_of(vars_to_keep)),
  bop_dec_2019 %>% select(any_of(vars_to_keep)),
  bop_nov_2018 %>% select(any_of(vars_to_keep)),
  bop_may_2018 %>% select(any_of(vars_to_keep)),
  bop_nov_2017 %>% select(any_of(vars_to_keep))
)

```

```{r}
dk_nr_cols <- bop_panel %>%
  select(where(is.factor)) %>%
  map_lgl(~ any(c("DK", "NR") %in% levels(.x))) %>%
  which() %>%
  names()

bop_panel <- bop_panel %>%
  mutate(across(all_of(dk_nr_cols), ~ fct_collapse(.x, "DK/NR" = c("DK", "NR"))))

bop_panel %>%
  select(where(is.factor)) %>%
  map(~ any(c("DK", "NR") %in% levels(.x))) %>%
  keep(~ .x == TRUE) %>%
  names()
```

# Analysis

## Direction
```{r}
direction_df <- bop_panel %>%
  group_by(wave, date) %>%
  summarise(
    pct_bad = mean(direction == "Bad", na.rm = TRUE),
    pct_good = mean(direction == "Good", na.rm = TRUE),
  )
```

```{r}
ggplot(direction_df, aes(x = date, y = pct_bad)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage (Bad)")
```

### Macro-Region
```{r}
direction_region_df <- bop_panel %>%
  group_by(wave, date, Macro_Region) %>%
  summarise(
    pct_bad = mean(direction == "Bad", na.rm = TRUE),
    pct_good = mean(direction == "Good", na.rm = TRUE),
  )

direction_region_df <- remove_missing(direction_region_df)
```

```{r}
ggplot(direction_region_df, aes(x = date, y = pct_bad)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~Macro_Region) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage (Bad)")
```

### Ethnicity
```{r}
direction_ethnicity_df <- bop_panel %>%
  group_by(wave, date, Ethnicity) %>%
  summarise(
    pct_bad = mean(direction == "Bad", na.rm = TRUE),
    pct_good = mean(direction == "Good", na.rm = TRUE),
  )

direction_ethnicity_df <- remove_missing(direction_ethnicity_df)

 direction_ethnicity_df <- direction_ethnicity_df %>%
   filter(Ethnicity != "DK/NR") 
```

```{r}
ggplot(direction_ethnicity_df, aes(x = date, y = pct_bad)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~Ethnicity) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Proportion (Bad)")
```

### Language
```{r}
direction_language_df <- bop_panel %>%
  group_by(wave, date, Language_Questionnaire) %>%
  summarise(
    pct_bad = mean(direction == "Bad", na.rm = TRUE),
    pct_good = mean(direction == "Good", na.rm = TRUE),
  )

direction_language_df <- remove_missing(direction_language_df)

 direction_language_df <- direction_language_df %>%
   filter(Language_Questionnaire != "DK/NR") 
```

```{R}
ggplot(direction_language_df, aes(x = date, y = pct_bad)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~Language_Questionnaire) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage (Bad)")
```

## Problems
```{r}
# mentioning problems across the 1st/2nd/3rd pref
problems_long <- bop_panel %>%
  select(wave, date, id, First_Problem, Second_Problem, Third_Problem) %>%
  pivot_longer(
    cols=c(First_Problem, Second_Problem, Third_Problem),
    names_to = "rank",
    values_to = "problem"
  )
```

```{r}
top_problems <- problems_long %>%
  filter(!is.na(problem)) %>%
  group_by(wave, date, problem) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(prop = n/sum(n))
```

```{r}
top_problems %>%
  filter(problem == "Economy" | problem == "Corruption" | problem == "Standard of living" | problem == "Internal order" | problem == "Corruption" | problem == "Independent Judiciary" | problem == "Peace and understanding among people") %>%
  ggplot(aes(x = date, y = prop)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~problem) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

```{r}
# most cited first problem
first_problem <- bop_panel %>%
  group_by(wave, date, First_Problem) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n))

first_problem <- remove_missing(first_problem)
```

```{r}
first_problem %>%
  filter(First_Problem == "Economy" | First_Problem == "Corruption" | First_Problem == "Standard of living" | First_Problem == "Internal order" | First_Problem == "Corruption" | First_Problem == "Independent Judiciary" | First_Problem == "Peace and understanding among people") %>%
  ggplot(aes(x = date, y = percentage)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~First_Problem) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

## Media

```{r}
bop_panel$Trust_Media <- fct_collapse(bop_panel$Trust_Media ,
                             "DK/NR"    = c("DK/NR", "None"),
                             "Internet" = c("Internet", "Social media"))
bop_panel$Important_Media <- fct_collapse(bop_panel$Important_Media ,
                             "DK/NR"    = c("DK/NR", "None"),
                             "Internet" = c("Internet", "Social media"))
bop_panel$Second_Media <- fct_collapse(bop_panel$Second_Media ,
                             "DK/NR"    = c("DK/NR", "None"),
                             "Internet" = c("Internet", "Social media"))
```

### Most trusted
```{r}
trusted_media <- bop_panel %>%
  group_by(wave, date, Trust_Media) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n))

trusted_media <- remove_missing(trusted_media)
```

```{r}
trusted_media %>%
  filter(Trust_Media == "Internet" | Trust_Media == "TV" | Trust_Media == "Radio" | Trust_Media == "Newspapers" | Trust_Media == "Family" | Trust_Media == "Friends and neighbours") %>%
  ggplot(aes(x = date, y = percentage)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~Trust_Media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

#### Language
```{r}
# most trusted form of media by language of survey

trusted_media_lang <- bop_panel %>%
  group_by(wave, date, Trust_Media, Language_Questionnaire) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n))

trusted_media_lang <- remove_missing(trusted_media_lang)
```

```{R}
trusted_media_lang %>%
  filter(Trust_Media == "Internet" | Trust_Media == "TV" | Trust_Media == "Radio" | Trust_Media == "Newspapers" | Trust_Media == "Family" | Trust_Media == "Friends and neighbours" | Trust_Media == "DK/NR") %>%
  ggplot(aes(x = date, y = percentage, color = Language_Questionnaire)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~Trust_Media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

#### Age
```{r}
# most trusted form of media by language of survey

trusted_media_age <- bop_panel %>%
  group_by(wave, date, Trust_Media, Age.f) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n))

trusted_media_age <- remove_missing(trusted_media_age)
```

```{R}
trusted_media_age %>%
  filter(Trust_Media == "Internet" | Trust_Media == "TV" | Trust_Media == "Radio" | Trust_Media == "Newspapers" | Trust_Media == "Family" | Trust_Media == "Friends and neighbours") %>%
  ggplot(aes(x = date, y = percentage, color = Age.f)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~Trust_Media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

### Most important media
```{r}
# important media mentioned across the 1st/2nd pref
media_pref_long <- bop_panel %>%
  select(wave, date, id, Important_Media, Second_Media) %>%
  pivot_longer(
    cols=c(Important_Media, Second_Media),
    names_to = "rank",
    values_to = "media"
  )

top_media_pref <- media_pref_long %>%
  filter(!is.na(media)) %>%
  group_by(wave, date, media) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(prop = n/sum(n))
```

```{r}
top_media_pref %>%
  filter(media == "Internet" | media == "TV" | media == "Radio" | media == "Newspapers" | media == "Family" | media == "Friends and neighbours") %>%
  ggplot(aes(x = date, y = prop)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```
#### Language
```{r}
media_pref_long_lang <- bop_panel %>%
  select(wave, date, id, Language_Questionnaire, Important_Media, Second_Media) %>%
  pivot_longer(
    cols=c(Important_Media, Second_Media),
    names_to = "rank",
    values_to = "media"
  )

top_media_pref_lang <- media_pref_long_lang %>%
  filter(!is.na(media)) %>%
  group_by(wave, date, Language_Questionnaire, media) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(prop = n/sum(n))

top_media_pref_lang <- remove_missing(top_media_pref_lang)
```

```{r}
top_media_pref_lang %>%
  filter(media == "Internet" | media == "TV" | media == "Radio" | media == "Newspapers" | media == "Family" | media == "Friends and neighbours") %>%
  ggplot(aes(x = date, y = prop, color = Language_Questionnaire)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")

```

#### Education
```{r}
media_pref_long_edu <- bop_panel %>%
  select(wave, date, id, Education, Important_Media, Second_Media) %>%
  pivot_longer(
    cols=c(Important_Media, Second_Media),
    names_to = "rank",
    values_to = "media"
  )

top_media_pref_edu <- media_pref_long_edu %>%
  filter(!is.na(media)) %>%
  group_by(wave, date, Education, media) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(prop = n/sum(n))

top_media_pref_edu <- remove_missing(top_media_pref_edu)
```

```{r}
top_media_pref_edu %>%
  filter(media == "Internet" | media == "TV" | media == "Radio" | media == "Newspapers" | media == "Family" | media == "Friends and neighbours") %>%
  ggplot(aes(x = date, y = prop, color = Education)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~media) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Date",
       y = "Percentage")
```

## Domestic politics
```{r}
bop_panel$year <- year(bop_panel$date)
bop_panel$year.f <- as.factor(bop_panel$year)
bop_panel$wave <- as.factor(bop_panel$wave)

bop_panel$Sandu <- as.numeric(bop_panel$Sandu)
bop_panel$Dodon <- as.numeric(bop_panel$Dodon)
bop_panel$Sor <- as.numeric(bop_panel$Sor)
bop_panel$PAS <- as.numeric(bop_panel$PAS)
bop_panel$PSRM <- as.numeric(bop_panel$PSRM)
```

```{r}
domestic_bop <- bop_panel %>%
  group_by(wave, date) %>%
  summarise(
    Sandu_mean = mean(Sandu, na.rm = TRUE),
    Sandu_sd = sd(Sandu, na.rm = TRUE),
    Sandu_n = sum(!is.na(Sandu)),
    Sandu_se = Sandu_sd / sqrt(Sandu_n),
    Sandu_lower = Sandu_mean - 1.96 * Sandu_se,
    Sandu_upper = Sandu_mean + 1.96 * Sandu_se,
    Dodon_mean = mean(Dodon, na.rm = TRUE),
    Dodon_sd = sd(Dodon, na.rm = TRUE),
    Dodon_n = sum(!is.na(Dodon)),
    Dodon_se = Dodon_sd / sqrt(Dodon_n),
    Dodon_lower = Dodon_mean - 1.96 * Dodon_se,
    Dodon_upper = Dodon_mean + 1.96 * Dodon_se,
    Sor_mean = mean(Sor, na.rm = TRUE),
    Sor_sd = sd(Sor, na.rm = TRUE),
    Sor_n = sum(!is.na(Sor)),
    Sor_se = Sor_sd / sqrt(Sor_n),
    Sor_lower = Sor_mean - 1.96 * Sor_se,
    Sor_upper = Sor_mean + 1.96 * Sor_se  
  )
```

```{r}
domestic_bop_long <- domestic_bop %>%
  pivot_longer(
    cols = c(Sandu_mean, Dodon_mean, Sor_mean, Sandu_lower, Sandu_upper, Dodon_lower, Dodon_upper, Sor_lower, Sor_upper),
    names_to = c("politician", ".value"),
    names_pattern = "(Sandu|Dodon|Sor)_(.*)"
  )
```

```{r}
ggplot(domestic_bop_long, aes(x = date, y = mean, color = politician, group = politician)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = politician), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  scale_y_reverse(limits = c(1,4)) +
  geom_point(size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Age
```{r}
domestic_bop_age <- bop_panel %>%
  group_by(wave, date, Age.f) %>%
  summarise(
    Sandu_mean = mean(Sandu, na.rm = TRUE),
    Sandu_sd = sd(Sandu, na.rm = TRUE),
    Sandu_n = sum(!is.na(Sandu)),
    Sandu_se = Sandu_sd / sqrt(Sandu_n),
    Sandu_lower = Sandu_mean - 1.96 * Sandu_se,
    Sandu_upper = Sandu_mean + 1.96 * Sandu_se,
    Dodon_mean = mean(Dodon, na.rm = TRUE),
    Dodon_sd = sd(Dodon, na.rm = TRUE),
    Dodon_n = sum(!is.na(Dodon)),
    Dodon_se = Dodon_sd / sqrt(Dodon_n),
    Dodon_lower = Dodon_mean - 1.96 * Dodon_se,
    Dodon_upper = Dodon_mean + 1.96 * Dodon_se,
    Sor_mean = mean(Sor, na.rm = TRUE),
    Sor_sd = sd(Sor, na.rm = TRUE),
    Sor_n = sum(!is.na(Sor)),
    Sor_se = Sor_sd / sqrt(Sor_n),
    Sor_lower = Sor_mean - 1.96 * Sor_se,
    Sor_upper = Sor_mean + 1.96 * Sor_se  
  )
```

```{r}
domestic_bop_age_long <- domestic_bop_age %>%
  pivot_longer(
    cols = c(Sandu_mean, Dodon_mean, Sor_mean, Sandu_lower, Sandu_upper, Dodon_lower, Dodon_upper, Sor_lower, Sor_upper),
    names_to = c("politician", ".value"),
    names_pattern = "(Sandu|Dodon|Sor)_(.*)"
  )
```

```{r}
ggplot(domestic_bop_age_long, aes(x = date, y = mean, color = politician, group = politician)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = politician), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  scale_y_reverse(limits = c(1,4)) +
  geom_point(size = 2) +
  facet_wrap(~Age.f) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Language
```{r}
domestic_bop_language <- bop_panel %>%
  group_by(wave, date, Language_Questionnaire) %>%
  summarise(
    Sandu_mean = mean(Sandu, na.rm = TRUE),
    Sandu_sd = sd(Sandu, na.rm = TRUE),
    Sandu_n = sum(!is.na(Sandu)),
    Sandu_se = Sandu_sd / sqrt(Sandu_n),
    Sandu_lower = Sandu_mean - 1.96 * Sandu_se,
    Sandu_upper = Sandu_mean + 1.96 * Sandu_se,
    Dodon_mean = mean(Dodon, na.rm = TRUE),
    Dodon_sd = sd(Dodon, na.rm = TRUE),
    Dodon_n = sum(!is.na(Dodon)),
    Dodon_se = Dodon_sd / sqrt(Dodon_n),
    Dodon_lower = Dodon_mean - 1.96 * Dodon_se,
    Dodon_upper = Dodon_mean + 1.96 * Dodon_se,
    Sor_mean = mean(Sor, na.rm = TRUE),
    Sor_sd = sd(Sor, na.rm = TRUE),
    Sor_n = sum(!is.na(Sor)),
    Sor_se = Sor_sd / sqrt(Sor_n),
    Sor_lower = Sor_mean - 1.96 * Sor_se,
    Sor_upper = Sor_mean + 1.96 * Sor_se  
  )

domestic_bop_language <- remove_missing(domestic_bop_language)
```

```{r}
domestic_bop_language_long <- domestic_bop_language %>%
  pivot_longer(
    cols = c(Sandu_mean, Dodon_mean, Sor_mean, Sandu_lower, Sandu_upper, Dodon_lower, Dodon_upper, Sor_lower, Sor_upper),
    names_to = c("politician", ".value"),
    names_pattern = "(Sandu|Dodon|Sor)_(.*)"
  )
```

```{r}
ggplot(domestic_bop_language_long, aes(x = date, y = mean, color = politician, group = politician)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = politician), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  scale_y_reverse(limits = c(1,4)) +
  geom_point(size = 2) +
  facet_wrap(~Language_Questionnaire) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Regression analysis
```{r}

Sandu_models <- list(
  "Age" = lm(Sandu ~ Age, data = bop_panel),
  "Gender" = lm(Sandu ~ Gender, data = bop_panel),
  "Ethnicity" = lm(Sandu ~ Ethnicity, data = bop_panel),
  "Language" = lm(Sandu ~ Language_Questionnaire, data = bop_panel),
  "Ecology" = lm(Sandu ~ Ecology, data = bop_panel),
  "Education" = lm(Sandu ~ Education, data = bop_panel),
  "Socioeconomic" = lm(Sandu ~ Socioeconomic, data = bop_panel),
  "Macro Region" = lm(Sandu ~ Macro_Region, data = bop_panel),
  "Most Trusted Media" = lm(Sandu ~ Trust_Media, data = bop_panel),
  "Direction" = lm(Sandu ~ direction, data = bop_panel),
  "Year" = lm(Sandu ~ year, data = bop_panel),
  "Unification" = lm(Sandu ~ Unification_For, data = bop_panel),
  "All" = lm(Sandu ~ Age + Gender + Ethnicity + Language_Questionnaire + Ecology + Education + Socioeconomic + Macro_Region + Trust_Media + Unification_For + direction + year, data = bop_panel)
)

modelsummary(Sandu_models, stars = TRUE)
```

```{r}
Sor_models <- list(
  "Age" = lm(Sor ~ Age, data = bop_panel),
  "Gender" = lm(Sor ~ Gender, data = bop_panel),
  "Ethnicity" = lm(Sor ~ Ethnicity, data = bop_panel),
  "Language" = lm(Sor ~ Language_Questionnaire, data = bop_panel),
  "Ecology" = lm(Sor ~ Ecology, data = bop_panel),
  "Education" = lm(Sor ~ Education, data = bop_panel),
  "Socioeconomic" = lm(Sor ~ Socioeconomic, data = bop_panel),
  "Macro Region" = lm(Sor ~ Macro_Region, data = bop_panel),
  "Wave" = lm(Sor ~ wave, data = bop_panel),
  "All" = lm(Sor ~ Age + Gender + Ethnicity + Language_Questionnaire + Ecology + Education + Socioeconomic + Macro_Region + wave, data = bop_panel)
)

modelsummary(Sor_models, stars = TRUE)
```

```{r}
Dodon_models <- list(
  "Age" = lm(Dodon ~ Age, data = bop_panel),
  "Gender" = lm(Dodon ~ Gender, data = bop_panel),
  "Ethnicity" = lm(Dodon ~ Ethnicity, data = bop_panel),
  "Language" = lm(Dodon ~ Language_Questionnaire, data = bop_panel),
  "Ecology" = lm(Dodon ~ Ecology, data = bop_panel),
  "Education" = lm(Dodon ~ Education, data = bop_panel),
  "Socioeconomic" = lm(Dodon ~ Socioeconomic, data = bop_panel),
  "Macro Region" = lm(Dodon ~ Macro_Region, data = bop_panel),
  "Wave" = lm(Dodon ~ wave, data = bop_panel),
  "All" = lm(Dodon ~ Age + Gender + Ethnicity + Language_Questionnaire + Ecology + Education + Socioeconomic + Macro_Region + wave, data = bop_panel)
)

modelsummary(Dodon_models, stars = TRUE)
```

# International politics
```{r}
bop_panel$Unification_For <- ifelse(bop_panel$Unification == "For", 1, 0)
bop_panel$Nato_For <- ifelse(bop_panel$NATO == "For", 1, 0)
bop_panel$EU_For <- ifelse(bop_panel$EU == "For", 1, 0)
bop_panel$Eurasia_For <- ifelse(bop_panel$Eurasia == "For", 1, 0)
bop_panel$EU_VS <- ifelse(bop_panel$EU_vs_Eurasia == "EU", 1, 0)
bop_panel$Eurasia_VS <- ifelse(bop_panel$EU_vs_Eurasia == "Eurasia", 1, 0)
```

## Unification
```{r}
unification_support <- bop_panel %>%
  group_by(wave, date, Unification_For) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n))

unification_support <- remove_missing(unification_support)
```

```{r}
unification_support %>%
  filter(Unification_For == 1) %>%
  ggplot(aes(x = date, y = percentage)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
Unification_models <- list(
  "Age" = glm(Unification_For ~ Age, data = bop_panel, family = "binomial"),
  "Gender" = glm(Unification_For ~ Gender, data = bop_panel, family = "binomial"),
  "Ethnicity" = glm(Unification_For ~ Ethnicity, data = bop_panel, family = "binomial"),
  "Language" = glm(Unification_For ~ Language_Questionnaire, data = bop_panel, family = "binomial"),
  "Ecology" = glm(Unification_For ~ Ecology, data = bop_panel, family = "binomial"),
  "Education" = glm(Unification_For ~ Education, data = bop_panel, family = "binomial"),
  "Socioeconomic" = glm(Unification_For ~ Socioeconomic, data = bop_panel, family = "binomial"),
  "Macro Region" = glm(Unification_For ~ Macro_Region, data = bop_panel, family = "binomial"),
  "Year" = glm(Unification_For ~ year, data = bop_panel, family = "binomial"),
  "Nato" = glm(Unification_For ~ Nato_For, data = bop_panel, family = "binomial"),
  "EU" = glm(Unification_For ~ EU_For, data = bop_panel, family = "binomial"),
  "All" = glm(Unification_For ~ Age + Gender + Ethnicity + Language_Questionnaire + Ecology + Education + Socioeconomic + Macro_Region + Nato_For + EU_For + year, data = bop_panel, family = "binomial")
)

modelsummary(Unification_models, stars = TRUE)
```

## EU support
```{r}
# multilateral vs unilateral support
EU_support <- bop_panel %>%
  group_by(wave, date, EU_For) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n)) %>%
  filter(EU_For == 1)

EU_support$Variable <- "EU support (not binary)"

EU_support$EU_For <- NULL

EU_support2 <- bop_panel %>%
  group_by(wave, date, EU_VS) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n)) %>%
  filter(EU_VS == 1)

EU_support2$EU_VS <- NULL

EU_support2$Variable <- "EU support (binary)"

EU_support <- rbind(EU_support, EU_support2)
EU_support$IO <- "EU"
```

```{r}
ggplot(EU_support, aes(x = date, y = percentage, color = Variable)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## NATO support
```{r}
NATO_support <- bop_panel %>%
  group_by(wave, date, Nato_For) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n)) %>%
  filter(Nato_For == 1)
```

```{r}
ggplot(NATO_support, aes(x = date, y = percentage)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Eurasia support
```{r}
Eurasia_support <- bop_panel %>%
  group_by(wave, date, Eurasia_For) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n)) %>%
  filter(Eurasia_For == 1)

Eurasia_support$Variable <- "Eurasia support (not binary)"

Eurasia_support$Eurasia_For <- NULL

Eurasia_support2 <- bop_panel %>%
  group_by(wave, date, Eurasia_VS) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(wave, date) %>%
  mutate(percentage = n/sum(n)) %>%
  filter(Eurasia_VS == 1)

Eurasia_support2$Eurasia_VS <- NULL

Eurasia_support2$Variable <- "Eurasia support (binary)"

Eurasia_support <- rbind(Eurasia_support, Eurasia_support2)

Eurasia_support$IO <- "Eurasia"
```

```{r}
ggplot(Eurasia_support, aes(x = date, y = percentage, color = Variable)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
IO_support <- rbind(EU_support, Eurasia_support)

ggplot(IO_support, aes(x = date, y = percentage, color = Variable)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~IO) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```